Localisation and Sensor Privacy Using the Extended Information Filter and Private Linear-Combination Aggregation\thanksref{footnoteinfo}

### 2020-09-03 main.tex
Fixed some things in problem formulation notes and from filter description yesterday - also added the fact that we only consider 2 dimensional localisation

The problem we will consider is a model-based localisation estimation problem in the presence of $n$ range-only sensors. This is defined by the typical process model
\begin{equation}
    x_k = f(x_{k-1})+w_k
\end{equation}
and measurement model
\begin{equation}
    z_k = h(x_k)+v_k\,,
\end{equation}
at time $k$, with zero-mean Gaussian noises $w_k \sim \mathcal{N}(0, Q)$ and $v_k \sim \mathcal{N}(0, R)$. The measurement function $h(\cdot)$ for range-sensors is dependant on state location vector $x_{k,loc}$ and sensor location $s_{i,loc},\,1\leq i \leq n$. It is given by
\begin{equation}
    h(x_k) = \|x_{k,loc} - s_{i,loc}\|\,.
\end{equation}
We are interested in running the above estimation filter in which the estimator (in this case navigator) does not disclose state information $x_k$ to sensors, while sensors do not disclose their individual measurements $z_k$, and locations $s_{i,loc}$, to the navigator.

We motivate this scenario with the example of aircraft navigation, when in the presence of privately-owned range-measuring towers. It is reasonable to assume that the current state of an aircraft may wish to be kept private from unknown tower-owning parties, while sensor tower locations may not wish to be disclosed to any navigating aircraft.

@@

Prediction step can be computed in either the EIF form or the equivalent EKF state and covariance can be computed and the prediction computed normally

@@

Using the EIF and distance sensors, we see that the required computation of measurement vector and matrices is reduced to computing $H^\top R^{-1}(z - h(x) + Hx)$ and $H^\top R^{-1}H$, for one-dimensional $R$ and jacobian $H$, at each sensor. As this is requires current state information $x$, relevant attributes need to be sent to the sensors.

We wish to encrypt current state information and use the encryption scheme defined to compute measurement information $H^\top R^{-1}(z - h(x) + Hx)$ and $H^\top R^{-1}H$ as a linear combination of state information, and encrypt it for aggregation. Thus allowing the navigator to only learn sums $\sum^n_{i=1}H^\top R^{-1}_i(z_i - h(x) + Hx)$ and $\sum^n_{i=1}H^\top R_i^{-1}H$ and update its position.

@@

The requirement introduced above, is that $H^\top R^{-1}(z - h(x) + Hx)$ and $H^\top R^{-1}H$ can be computed as linear combinations of state-derived variables. Here $H$ is a Jacobian and includes current state information as well.

State-derived variables refer to state vector elements, as well as non-linear functions of these elements, which do not require sensor position or measurement information to be computed. As our encryption scheme described above, only supports linear combinations of weights, in this case state-derived variables, all non-linear functions required for measurements information to be computed, need to be treated as individual weights to be combined.

From the definition of matrix multiplication, we see that any $KxL$ matrix $D=ABC$, where $B$ has dimensions $MxN$, has element $d_{i,j}$ defined by
\begin{equation*}
    d_{i,j} = \sum^{N}_{n=1}\sum^{M}_{m=1}a_{i,m}b_{m,n}c_{n,j}\,,
\end{equation*}
and can therefore be considered as a linear combination of all elements $a_{i,m}c_{n,j}$. Applying this to the measurement matrix and vector $H^\top R^{-1}(z - h(x) + Hx)$ and $H^\top R^{-1}H$, we require that all values of $h(x)$, and all elements in Jacobian $H$ be linear combinations of state-derived variables

Next, we show how this can be achieved for the specific case of 2-dimensional localisation, with range-only sensors.

@@

\subsection{Localisation Measurement Model}
% Show here the weighted integrals that give mean and variance of the new noise. If wanting to show more working, do this in the appendix section, but probably not needed.

@@

From our problem statements in Section $\cdot$, we specify the goal of private localisation in the 2-dimensional case, using range-only sensors. Given the current location estimate $x_{loc} = (x,y)$, which may comprise only part of the filter state estimate, and sensor location $s_{loc} = (s_x, s_y)$, the measurement equation is given by
\begin{equation*}
    \begin{split}
        h(x) &= \|x_{loc} - s_{loc}\| \\
        &= \sqrt{(x - s_x)^2 + (y - s_y)^2}\,.
    \end{split}
\end{equation*}
The Jacobian $H$ for the measurement function above at $x_{loc,0}=(x_0, y_0)$ is then given by
\begin{equation*}
    H = 
    \begin{bmatrix}
        \frac{x_0 - s_x}{\sqrt{(x_0 - s_x)^2 + (y_0 - s_y)^2}} \\
        \frac{y_0 - s_y}{\sqrt{(x_0 - s_x)^2 + (y_0 - s_y)^2}} 
    \end{bmatrix}\,.
\end{equation*}
The problem with the above function $h$ and Jacobian $H$ is that they cannot be reformulated as a linear combination of weights, where weights are defined by functions of $x$ and $y$, and do not include $s_x$ and $s_y$.

Our proposed solution to this problem, is working with the adjust measurement function $h'(x) = h(x)^2$. The method for modifying a normal range-sensor measurement in this way and handling the no-longer Gaussian measurement noise with variance dependant on distance, will be discussed in Section \ref{subsec:range_measurement_mod}. While this transformation complicated the measurement model noise, it allows for the linear formulation required for the encryption scheme, and the privacy of $H^{\prime\top} R^{\prime -1}(z' - h(x) + Hx)$ and $H^{\prime\top} R^{\prime -1}H'$. The adjusted measurement function is now given by
\begin{equation*}
    \begin{split}
        h'(x) &= \left(\|x_{loc} - s_{loc}\|\right)^2 \\
        &= (x - s_x)^2 + (y - s_y)^2 \\
        &= x^2 + y^2 -2s_xx -2s_yy +s_x^2 +s_y^2
    \end{split}\,,
\end{equation*}
and Jacobian $H'$ by
\begin{equation*}
    H' = 
    \begin{bmatrix}
        2x_0 - 2s_x \\
        2y_0 - 2s_y
    \end{bmatrix}\,.
\end{equation*}
The above function and jacobian elements are linear combinations of state-derived functions $x$, $x^2$, $y$, $xy$ and $y^2$. From our equation $\cdot$ in the previous section, we can now define $H^{\prime\top} R^{\prime -1}(z' - h(x) + Hx)$ and $H^{\prime\top} R^{\prime -1}H'$ in terms of state-derived functions $x$, $x^2$, $x^3$, $y$, $xy$, $x^2y$, $y^2$, $xy^2$ and $y^3$.
\begin{equation*}
    \begin{split}
        H&^{\prime\top} R^{\prime -1}(z' - h(x) + Hx) = \\
        &\begin{bmatrix}
            \begin{split}
                &(2R^{\prime -1}z')x + (-2R^{\prime -1}s_xz) + (-2R^{\prime -1}s_x^2)x \\
                &\quad+ (2R^{\prime -1}s_x^3) + (-2R^{\prime -1}s_y^2)x + (2R^{\prime -1}s_xs_y^2) \\
                &\quad+ (2R^{\prime -1})x^3 + (-R^{\prime -1}s_x)x^2 + (2R^{\prime -1})xy^2 \\
                &\quad+ (-2R^{\prime -1}s_x)y^2
            \end{split}
            \vspace{5pt}\\
            \begin{split}
                &(2R^{\prime -1}z)y + (-2R^{\prime -1}s_yz) + (-2R^{\prime -1}s_x^2)y \\
                &\quad+ (2R^{\prime -1}s_ys_x^2) + (-2R^{\prime -1}s_y^2)y + (2R^{\prime -1}s_y^3) \\
                &\quad+ (2R^{\prime -1})x^2y + (-2R^{\prime -1}s_y)x^2 + (2R^{\prime -1})y^3 \\
                &\quad+ (-2R^{\prime -1}s_y)y^2
            \end{split}
        \end{bmatrix}
    \end{split}
\end{equation*}
and
\begin{equation*}
    H^{\prime\top} R^{\prime -1}H' = 
    \begin{bmatrix}
        \alpha_{11} & \alpha_{12} \\
        \alpha_{21} & \alpha_{22}
    \end{bmatrix}\,,
\end{equation*}
with
\begin{align*}
    \alpha_{11} &= (4R^{\prime -1})x^2 + (-8R^{\prime -1}s_x)x + (4R^{\prime -1}s_x^2) \\
    \begin{split}
        \alpha_{12} &= (4R^{\prime -1})xy + (-4R^{\prime -1}s_y)x + (-4R^{\prime -1}s_x)y \\
        &\qquad+ (4R^{\prime -1}s_xs_y)
    \end{split}\\
    \alpha_{21} &= \alpha_{12} \\
    \alpha_{22} &= (4R^{\prime -1})y^2 + (-8R^{\prime -1}s_y)y + (4R^{\prime -1}s_y^2)\,.
\end{align*}

@@

Lastly, since we aggregate multiple linear combinations at each time step $k$ (each element in the result matrix and vector comes from a separate aggregation), we require unique hash values for each aggregation. This is handled by hashing by using setting time variable $t$ as concatenation $k \| i \| j$ for each element $x_{i,j}$. Aggregation of each element is now performed under a different encryption scheme ``time'', when computing \eqref{eqn:our_scheme_lin_comb}.

@@

\subsection{Range Measurement Modification} \label{subsec:range_measurement_mod}
The conversion of measurement function $h'(x) = h(x)^2$ is in practice complicated by gausian nature of the range sensor measurement noise. As we only know $z = h(x) + n$, where $n~\mathcal{N}(0,\sigma)$, we can estimate the new measurement $z'= z^2$. This results in the new measurement
\begin{equation*}
    \begin{split}
        z^2 &= (h(x) + n)^2 \\
        &= h(x)^2 + 2nh(x) + n^2 \\
        &= h'(x) + 2nh(x) + n^2
    \end{split}\,,
\end{equation*}
where noise term $2nh(x) + n^2$ has mean $\sigma^2$ and standard deviation $2(2h(x)^2\sigma^2 + \sigma^4)$. We account for the non-zero mean by offsetting the squared measurement and taking this as the adjusted measurement model
\begin{equation*}
    \begin{split}
        z' &= h'(x) + 2nh(x) + n^2 - \sigma^2 \\
        &= h'(x) + n'
    \end{split}\,,
\end{equation*}
with variance of $n'$ approximated as $2(2z^2\sigma^2 + \sigma^4)$, by using the original measurement $z$ as an unbiased estimate for $h(x)$.

@@

\subsection{Leakage} \label{subsec:leakage}
We recall the assumptions made in problem formulation and our encryption scheme, and the allowed leakage of the encryption scheme. Sensors are honest but curious, meaning sensor values cannot be modified from what is measured, and the protocol is followed exactly, and that the navigator sends the same state information to all sensors at each time step. This is justified to an extent with wireless broadcasting being received by all sensors in a typical localisation setting.

@@

The navigator, which may maliciously choose weights at every time step, is always restricted to learning at most the sum $\sum^n_{i=1} a_i$ for some values $a_i$ derived from sensor $i$ location and measurement. It is therefore intuitive to say that at most the navigator can learn the average of values kept private, that is, the navigator can at most learn the average position and location of sensors. The same can be said for any corrupted subset of honest but curious sensors and malicious navigator.

>>>>

The problem we will consider is a two-dimensional model-based localisation estimation problem in the presence of $n$ range-only sensors. This is defined by the typical process model
\begin{equation}
    x_k = f(x_{k-1})+w_k
\end{equation}
and measurement model
\begin{equation}
    z_k = h(x_k)+v_k\,,
\end{equation}
at time $k$, with zero-mean Gaussian noises $w_k \sim \mathcal{N}(0, Q)$ and $v_k \sim \mathcal{N}(0, R)$. The measurement function $h(\cdot)$ for range-sensors is dependant on state location vector $x_{k,loc}$ and sensor location $s_{i,loc},\,1\leq i \leq n$. It is given by
\begin{equation}
    h(x_k) = \|x_{k,loc} - s_{i,loc}\|\,.
\end{equation}
We are interested in running the above estimation filter in which the estimator (in this case navigator) does not disclose state information $x_k$ to sensors, while sensors do not disclose their individual measurements $z_k$, and locations $s_{i,loc}$, to the navigator.

We motivate this scenario with the example of aircraft navigation, when in the presence of privately-owned range-measuring towers. It is reasonable to assume that the current state of an aircraft may wish to be kept private from unknown tower-owning parties, while sensor tower locations may not wish to be disclosed to any navigating aircraft. Although we will consider the two-dimensional case for simplicity, the extension to three dimensions is trivial at the cost of communication complexity, as will be made clear in Section \ref{subsec:observation_model}.

@@

Prediction step can be computed in either the EIF form or the equivalent EKF state and covariance can be computed before performing the prediction step in EKF form and converting the result back.

@@

Using the EIF and distance sensors, we see that the required computation of measurement vector and matrices is reduced to computing $H^\top R^{-1}(z - h(x) + Hx)$ and $H^\top R^{-1}H$, for one-dimensional $R$ and jacobian $H$, at each sensor. As this requires current state information $x$, relevant attributes need to be sent to the sensors.

We wish to encrypt current state information and use the encryption scheme defined to compute measurement information $H^\top R^{-1}(z - h(x) + Hx)$ and $H^\top R^{-1}H$ as a linear combination of said state information, and encrypt it for aggregation. This then allows the navigator to learn only the sums $\sum^n_{i=1}H^\top R^{-1}_i(z_i - h(x) + Hx)$ and $\sum^n_{i=1}H^\top R_i^{-1}H$, which is sufficient to update its position.

@@

The requirement introduced above, is that measurement matrix and vector $H^\top R^{-1}(z - h(x) + Hx)$ and $H^\top R^{-1}H$ can be computed as linear combinations of functions involving only state vector elements. In particular, a linear combination of functions which depend only on information available to the navigator. In the measurement matrix and vector above, $H$ is the measurement function Jacobian, and includes current state information as well.

As our encryption scheme only supports linear combinations of weights, the weights need to capture all non-linear computations involving state vector elements. From the definition of matrix multiplication, we see that any $KxL$ matrix $D=ABC$, where $B$ has dimensions $MxN$, has element $d_{i,j}$ defined by
\begin{equation*}
    d_{i,j} = \sum^{N}_{n=1}\sum^{M}_{m=1}a_{i,m}b_{m,n}c_{n,j}\,,
\end{equation*}
and can therefore be considered as a linear combination of all elements $a_{i,m}c_{n,j}$. Applying this to the measurement matrix and vector $H^\top R^{-1}(z - h(x) + Hx)$ and $H^\top R^{-1}H$, we require that all values of $h(x)$, and all elements in Jacobian $H$ be linear combinations of functions dependant on state vector elements and nothing else.

We will show how this can be achieved for the specific case of 2-dimensional localisation, with range-only sensors.

@@

\subsection{Localisation Measurement Model} \label{subsec:observation_model}
% Show here the weighted integrals that give mean and variance of the new noise. If wanting to show more working, do this in the appendix section, but probably not needed.

@@

From our problem statement in Section \ref{sec:problem_statement}, we specify the goal of private localisation in the 2-dimensional case, using range-only sensors. Given the current location estimate $x_{loc} = (x,y)$, which may comprise only part of the filter state estimate, and sensor location $s_{loc} = (s_x, s_y)$, the measurement equation is given by
\begin{equation*}
    \begin{split}
        h(x) &= \|x_{loc} - s_{loc}\| \\
        &= \sqrt{(x - s_x)^2 + (y - s_y)^2}\,.
    \end{split}
\end{equation*}
The Jacobian $H$ for the measurement function above at $x_{loc,0}=(x_0, y_0)$ is then given by
\begin{equation*}
    H = 
    \begin{bmatrix}
        \frac{x_0 - s_x}{\sqrt{(x_0 - s_x)^2 + (y_0 - s_y)^2}} \\
        \frac{y_0 - s_y}{\sqrt{(x_0 - s_x)^2 + (y_0 - s_y)^2}} 
    \end{bmatrix}\,.
\end{equation*}
The problem with the above function $h$ and Jacobian $H$ is that they cannot be reformulated as a linear combination of weights, where weights do not depend on $s_x$ and $s_y$.

Our proposed solution to this problem, is working with the adjusted measurement function $h'(x) = h(x)^2$. The method for modifying a normal range-sensor measurement in this way and handling the no-longer Gaussian measurement noise with variance dependant on distance, will be discussed in Section \ref{subsec:range_measurement_mod}. While this transformation complicates the measurement model noise, it allows for the linear formulation required for the encryption scheme, and the privacy of $H^{\prime\top} R^{\prime -1}(z' - h(x) + Hx)$ and $H^{\prime\top} R^{\prime -1}H'$. The adjusted measurement function is now given by
\begin{equation*}
    \begin{split}
        h'(x) &= \left(\|x_{loc} - s_{loc}\|\right)^2 \\
        &= (x - s_x)^2 + (y - s_y)^2 \\
        &= x^2 + y^2 -2s_xx -2s_yy +s_x^2 +s_y^2
    \end{split}\,,
\end{equation*}
and Jacobian $H'$ by
\begin{equation*}
    H' = 
    \begin{bmatrix}
        2x_0 - 2s_x \\
        2y_0 - 2s_y
    \end{bmatrix}\,.
\end{equation*}
The above function and jacobian elements are linear combinations of functions $x$, $x^2$, $y$, $xy$ and $y^2$. From the previous section, we know we can now define $H^{\prime\top} R^{\prime -1}(z' - h(x) + Hx)$ and $H^{\prime\top} R^{\prime -1}H'$ in terms of more state vector dependant functions, in this case, $x$, $x^2$, $x^3$, $y$, $xy$, $x^2y$, $y^2$, $xy^2$ and $y^3$.
\begin{equation*}
    \begin{split}
        H&^{\prime\top} R^{\prime -1}(z' - h(x) + Hx) = \\
        &\begin{bmatrix}
            \begin{split}
                &(2R^{\prime -1}z')x + (-2R^{\prime -1}s_xz) + (-2R^{\prime -1}s_x^2)x \\
                &\quad+ (2R^{\prime -1}s_x^3) + (-2R^{\prime -1}s_y^2)x + (2R^{\prime -1}s_xs_y^2) \\
                &\quad+ (2R^{\prime -1})x^3 + (-R^{\prime -1}s_x)x^2 + (2R^{\prime -1})xy^2 \\
                &\quad+ (-2R^{\prime -1}s_x)y^2
            \end{split}
            \vspace{5pt}\\
            \begin{split}
                &(2R^{\prime -1}z)y + (-2R^{\prime -1}s_yz) + (-2R^{\prime -1}s_x^2)y \\
                &\quad+ (2R^{\prime -1}s_ys_x^2) + (-2R^{\prime -1}s_y^2)y + (2R^{\prime -1}s_y^3) \\
                &\quad+ (2R^{\prime -1})x^2y + (-2R^{\prime -1}s_y)x^2 + (2R^{\prime -1})y^3 \\
                &\quad+ (-2R^{\prime -1}s_y)y^2
            \end{split}
        \end{bmatrix}
    \end{split}
\end{equation*}
and
\begin{equation*}
    H^{\prime\top} R^{\prime -1}H' = 
    \begin{bmatrix}
        \alpha_{11} & \alpha_{12} \\
        \alpha_{21} & \alpha_{22}
    \end{bmatrix}\,,
\end{equation*}
with
\begin{align*}
    \alpha_{11} &= (4R^{\prime -1})x^2 + (-8R^{\prime -1}s_x)x + (4R^{\prime -1}s_x^2) \\
    \begin{split}
        \alpha_{12} &= (4R^{\prime -1})xy + (-4R^{\prime -1}s_y)x + (-4R^{\prime -1}s_x)y \\
        &\qquad+ (4R^{\prime -1}s_xs_y)
    \end{split}\\
    \alpha_{21} &= \alpha_{12} \\
    \alpha_{22} &= (4R^{\prime -1})y^2 + (-8R^{\prime -1}s_y)y + (4R^{\prime -1}s_y^2)\,.
\end{align*}

@@

Lastly, since we aggregate multiple linear combinations at each time step $k$ (each element in the result matrix and vector comes from a separate aggregation), we require unique hash values for each aggregation. This is handled by setting time variable $t$ as concatenation $k \| i \| j$ for each element $x_{i,j}$. Aggregation of each element is now performed under a different encryption scheme ``time'' step, when computing \eqref{eqn:our_scheme_lin_comb}.

The solution above has been derived for the two-dimensional localisation case but it is easy to see that a similar solution would work in three-dimensional space. The cost of increasing localisation dimension however, is the increased number of state vector element functions required for broadcast by the navigator, and length of linear combination computed by sensors.

@@

\subsection{Range Measurement Modification} \label{subsec:range_measurement_mod}
The conversion of measurement function $h'(x) = h(x)^2$ is in practice complicated by measurement noise manipulation. As we only know $z = h(x) + n$, where $n~\mathcal{N}(0,\sigma)$, we can estimate the new measurement as $z^2$. This results in the new measurement
\begin{equation*}
    \begin{split}
        z^2 &= (h(x) + n)^2 \\
        &= h(x)^2 + 2nh(x) + n^2 \\
        &= h'(x) + 2nh(x) + n^2
    \end{split}\,,
\end{equation*}
where noise term $2nh(x) + n^2$ has mean $\sigma^2$ and standard deviation $2(2h(x)^2\sigma^2 + \sigma^4)$. We account for the non-zero mean by offsetting the squared measurement and taking this as the adjusted measurement model
\begin{equation*}
    \begin{split}
        z' &= h'(x) + 2nh(x) + n^2 - \sigma^2 \\
        &= h'(x) + n'
    \end{split}\,,
\end{equation*}
with variance of $n'$ approximated as $2(2z^2\sigma^2 + \sigma^4)$, by using the original measurement $z$ as an unbiased estimate for $h(x)$.

@@

\subsection{Leakage} \label{subsec:leakage}
We recall the assumptions made in the problem formulation and our encryption scheme, and the allowed leakage of the encryption scheme. Sensors are honest but curious, meaning sensor values cannot be modified from what is measured, and the protocol is followed exactly, and that the navigator sends the same state information to all sensors at each time step. This is justified to an extent with wireless broadcasting being received by all sensors in a typical localisation setting.

@@

The navigator, which may maliciously choose weights at every time step, is always restricted to learning at most the sum $\sum^n_{i=1} a_i$ for some values $a_i$ derived from sensor $i$'s location and measurement. It is therefore intuitive to say that at most the navigator can learn the average of values kept private, that is, the navigator can at most learn the average position and location of sensors. The same can be said for any corrupted subset of honest but curious sensors and malicious navigator.
